Data Analysis Methods for the Detection of the Epoch of Reionization

### 2017-08-10 PAPER.tex
-Re-wrote signalloss explanation in toy model section

To build our intuition into how a particular choice of weighting, namely inverse covariance weighting, can lead to signal loss, we use a toy model. We construct a simple dataset that contains 2-dimensional data (representing visibility data with $100$ time integrations and $20$ frequency channels). This model represents realistic dimensions of about an hour of PAPER data which might be used for a power spectrum analysis. We create a mock bright foreground signal, $\textbf{x}_{FG}$, as a complex sinusoid that varies in time and frequency (\cc{explain why a sinusoid for FG}), as well as a mock EoR signal, $\textbf{x}_{EoR}$, as a complex, Gaussian random signal. Our combined data vector is $\textbf{x} = \textbf{x}_{FG} + \textbf{x}_{EoR}$, to which we apply different weighting schemes. The three data vectors are shown in Figure \ref{fig:toy_sigloss1}.

@@

One choice for a weight matrix $\textbf{R}$ is an inverse covariance matrix. This type of weighting is attractive for power spectrum analyses because it is an aggressive way to down-weight foregrounds and yields the smallest possible error bars on the measurements (\citealt{tegmark_et_al1997a}; \citealt{bond_et_al1998}). The covariance matrix $\textbf{C}$, which in our case describes covariances between frequency channels, can be estimated in a variety of ways. Given perfect foreground, instrumental, and EoR models, we could form $\textbf{C}$ in a way that accurately describes our measured data. However, if our foreground model is flawed for example, our estimate of $\textbf{C}$ would not be successful at down-weighting the foregrounds in the data.

@@

\noindent assuming $\langle\textbf{x}\rangle = 0$, where $\langle \rangle$ denotes a finite time-average. The inverse covariance matrix is therefore $\textbf{R} = \hat{\textbf{C}}^{-1}$.

@@

\begin{figure*}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 1cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss3.png}
	\includegraphics[trim={1cm 0.2cm 0cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss4.png}
	\caption{Resulting power spectrum estimates for the toy model simulation described in Section \ref{sec:toymodel} --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We use inverse covariance weighting where $\textbf{C}$ is derived from the data (left), contrasted with projecting out the first eigenmode only (right). There is signal loss for the former but not the latter. \cc{elaborate}}
	\label{fig:toy_sigloss3}
\end{figure*}

As shown, our inverse covariance-weighted result successfully suppresses foregrounds \cc{explain why}. It is also evident that our result fails to recover the EoR signal --- it exhibits the correct shape, but the amplitude level is slightly low. This is evidence of signal loss. In order to understand how this occurs, we need to closely study our covariance matrix, $\hat{\textbf{C}}$, which is shown in Figure \ref{fig:toy_sigloss12}.

@@

If $\textbf{C}$ is computed from the data itself and its inverse is then applied to the data, it carries the risk of over-fitting information in the data and introducing a multiplicative bias to estimates of the signal \cc{elaborate with math!}. As a contrasting example, if instead we choose $\textbf{R} = \textbf{I}$, there will be no signal loss (although it may suffer from foreground or instrumental bias).

It turns out that because we estimated $\textbf{C}$ from our data, its eigenspectrum offers insight into understanding how inverse covariance weighting affects our result. An eigenspectrum ranks the eigenvalues of $\hat{\textbf{C}}$ from highest to lowest and can be thought of as a spectrum of weights that are given to each frequency mode in the data. Each eigenvalue encodes the amount of variance in the dataset for a particular mode. The eigenspectrum of the identity matrix $\textbf{I}$ is flat (all $1$'s) because it gives equal weighting to all modes.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss2.png}
	\caption{Eigenspectrum of $\hat{\textbf{C}}_{FG}$ (blue), $\hat{\textbf{C}}_{EoR}$ (red), and $\hat{\textbf{C}}_{FG+EoR}$ (green). The eigenspectrum of $\hat{\textbf{C}}_{FG}$ peaks at the zeroth eigenmode, due to the presence of only one sinusoid. All eigenspectra steepen slightly for the final few eigenmodes, a behavior that results in some modes being upweighted more strongly than others, thereby overfitting noise and causing signal loss in the final power spectrum estimate (Figure \ref{fig:toy_sigloss3}, left). If only the first eigenmode is used, all others get equal weights and hence, there is no signal loss (Figure \ref{fig:toy_sigloss3}, right). \cc{also illustrate first few and last few eigenvalues -- how to best show this?}}
	\label{fig:toy_sigloss2}
\end{figure}

For our toy model, the eigenspectra of $\hat{\textbf{C}}_{FG}$, $\hat{\textbf{C}}_{EoR}$, and $\hat{\textbf{C}}$ are shown in Figure \ref{fig:toy_sigloss2}. The blue curve peaks only at the zeroth eigenmode, revealing that the foregrounds can be described by a single eigenmode. This is a direct consequence of how it was constructed (only one sinusoid). The eigenspectrum of EoR, however, is fairly flat, signifying that its covariance matrix is similar to an identity matrix. This makes sense because EoR consists of a random signal, which should only have strong frequency-frequency correlations when two frequencies are the same.

The eigenmodes with the highest eigenvalues will be down-weighted the most; when the spectrum is inverted, they will be given the smallest weights. In contrast, the modes with the smallest eigenvalues will be up-weighted the most. In other words, if an eigenspectrum is not perfectly flat, all modes will be given different weights and it is possible to up-weight a few modes much more than others. The subtle slope of the eigenspectrum in this toy example is the result of describing shapes in the data that aren't physically there \cc{reword sentence... the shape is there but it's not valid?}. This gives rise to a small amount of signal loss because it is possible to overfit a few modes (the highest weighted ones) ever so slightly. We will see in the next section how this effect can be exaggerated as the `steepness' of the spectrum changes.

Using what we've learned about the eigenspectrum, we can tweak it in a simple way to suppress foregrounds and yield zero signal loss. Namely, we can project out the zeroth eigenmode (i.e. zero out all eigenmodes except the first one), thereby down-weighting the foregrounds perfectly and nothing else. Altering $\hat{\textbf{C}}$ in this way is one example of a regularization method, in which we are changing $\hat{\textbf{C}}$ in a way that flattens its eigenspectrum, making it more identical to that of $\textbf{I}$. The resulting power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. In this case we recover EoR, demonstrating that if we can model our foregrounds, we can down-weight them without signal loss. There are several other ways to regularize $\hat{\textbf{C}}$, and we will discuss some in Section \ref{sec:otherweight}.

@@

We have shown how signal loss can arise due to unintentionally up-weighting eigenmodes with small eigenvalues (overfitting the noise). We will next show how this effect is exaggerated by reducing the total number of independent samples in a dataset.

@@

Because fringe-rate filtering is analogous to averaging in time, it comes at the cost of reducing the total number of independent samples in the data. To mimic this filter, we average every four time integrations of our toy model dataset together, yielding $25$ independent samples in time (Figure \ref{fig:toy_sigloss5}, left). We choose these numbers so that the total number of independent samples is similar to the number of frequency channels --- therefore, our matrices will be full rank. The resulting eigenspectra (Figure \ref{fig:toy_sigloss5}, right), as compared to those in Figure \ref{fig:toy_sigloss2}, fall more steeply for the last few eigenmodes. This slope is a consequence of reducing the variance of the dataset through averaging. More specifically, when numbers are averaged together we are absorbing outliers into the average, resulting in a narrower distribution of the data. Hence, the eigenvalues, which rank the strength of variances in the data, decreases more than an un-averaged case.

\begin{figure*}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss5.png}
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss6.png}
	\caption{Left: Our `fringe-rate filtered' (time-averaged) toy model dataset. We averaged every four samples together, yielding $25$ independent samples in time. Real parts are shown here. Right: Eigenspectrum of the covariance matrices of foregrounds only (blue), EoR only (red), and FG + EoR (green). The eigenspectrum of foregrounds only peaks at the zeroth eigenmode, a consequence of the presence of only one sinusoid. All eigenspectra steepen for the final few eigenmodes, a behavior that results in a few noisy modes being upweighted more strongly than others, thereby overfitting noise and causing signal loss in the final power spectrum estimate.}
	\label{fig:toy_sigloss5}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss7.png}
	\caption{Resulting power spectrum estimate for the fringe-rate filtered toy model simulation --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We use inverse covariance weighting where $\textbf{C}$ is derived from the data. There is a larger amount of signal loss than for un-fringe-rate filtered data, a consequence of a steeper eigenspectrum slope which results from washing out variances in the data through averaging. The power spectrum estimate is also noisier, since most of the information used in constructing the estimate came from only a few eigenmodes.}
	\label{fig:toy_sigloss7}
\end{figure}

The eigenspectrum shape has large consequences on the resulting power spectra, as evidenced by the power spectra shown in Figure \ref{fig:toy_sigloss7}. There is a much larger amount of signal loss for this time-averaged dataset. To reiterate, the drop-off in the eigenspectrum shape suggests that when the spectrum is inverted, a few modes are up-weighted much more strongly than the rest. \cc{math?} This dramatic up-weighting drowns out the EoR signal, giving rise to signal loss. Additionally, most of our power spectrum information is coming from only the last few modes and as a result, it is a noisier estimate. This is evident by noticing that the green curve in Figure \ref{fig:toy_sigloss7} fails to trace the shape of the unweighted EoR power spectrum.

Using our toy model, we have seen that a sensitivity-driven analysis technique like fringe-rate filtering has trade-offs of signal loss and noisier estimates. Longer integrations increase sensitivity but reduce the number of independent samples, resulting in steep drop-offs in eigenspectra.

@@

In Section \ref{sec:toymodel} we showed how altering $\hat{\textbf{C}}$ can make the difference between zero and some signal loss, if we can distinguish between foreground eigenmodes and EoR eigenmodes. We will now use our toy model to describe several other ways to tailor $\hat{\textbf{C}}$ in order to minimize signal loss. We illustrate the resulting power spectra and eigenspectra for four different cases in Figures \ref{fig:toy_sigloss8} and \ref{fig:toy_sigloss14}.

@@

The second panel (top right) in Figure \ref{fig:toy_sigloss8} uses a regularization method of setting $\hat{\textbf{C}} = \hat{\textbf{C}} + \textbf{I}$. By adding the identity, we are weighting the diagonal elements of the matrix more heavily than those off-diagonal, thereby flattening out its eigenspectrum \cc{why?}.

The third panel (bottom left) in Figure \ref{fig:toy_sigloss8} flattens out the eigenspectrum of $\hat{\textbf{C}}$ a different way - by zeroing all but the first three eigenmodes. This means that the first three eigenmodes will be given weights (the three highest variance modes will be down-weighted), but the rest will be treated with similar weights. Hence, we avoid up-weighting the last few noisy modes and have negligible signal loss. However, we do not perfectly recover the shape of EoR because we lost information when projecting out certain modes.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/data_errors.png}
	\caption{$2\sigma$ upper limit power spectrum estimates using PAPER-64 data and three different bootstrapping methods. The data is fringe-rate filtered and a weighting matrix of $\textbf{C}_{eff}^{-1}$ is used, where $\textbf{C}_{eff} = \textbf{C}*\textbf{I}$, in order to minimize signal loss. Three different bootstrapping methods are shown: the black and red use methods where bootstrapping occurs on the baseline axis only (not the time axis), and they differ by how the baselines are randomly sampled. It is evident that sensitivity is maximized when baselines are sampled in a way that ensures mostly independent samples (only the last baseline spot is sampled randomly). The blue result bootstraps over both the baseline and time axes, and illustrates how errors can be under-estimated if sampling more values than independent ones (fringe-rate filtering reduces the number of independent samples).}
	\label{fig:data_errors}
\end{figure}

@@

\noindent Because of the lower triangular matrix, this window function has the property of preventing the leakage of foreground power from low $k$ to high $k$ modes. Specifically, we order the elements in $\textbf{F}$ in such a way so that power can leak from high $k$ modes to low $k$ modes, but not vice versa. Since most foreground power shows up at low $k$'s, this method ensures a window function that retains clean, noise-dominated measurements while minimizing foreground bias.

@@

To mitigate crosstalk bias, we apply a fringe-rate filter that suppresses fringe-rates of zero... \cc{why don't we filter out fringe-rates of zero completely?}

@@

\appendix
\section{A toy model for signal loss}
\label{sec:SigLossToy}

@@

The minimum-variance quadratic estimator $\hat{p}_\alpha$ for the $\alpha$th bandpower of the power spectrum is given by \acl{We need to decide whether the Q matrices have their Greek indices as upper or lower indices.} \acl{Also need to state in a footnote that for this section only we'll assume without loss of generality that the data are real}
\begin{equation}
\hat{p}_\alpha = \frac{1} {2 \F_{\alpha \alpha} }\x^t \C^{-1} \Q^{\alpha} \C^{-1} \x,
\end{equation}
where
\begin{equation}
F_{\alpha \alpha} \equiv \frac{1}{2} \textrm{tr} \left( \C^{-1} \Q^\alpha \C^{-1} \Q^\alpha \right)
\end{equation}
is the $\alpha$th diagonal element of the Fisher matrix. In our case, however, we do not have \emph{a priori} knowledge of the covariance matrix. Thus, we replace $\C$ with $\Chat$, its data-derived approximation. Our estimator then becomes
\begin{equation}
\label{eq:phatloss}
\hat{p}_\alpha^\textrm{loss} = \frac{1} {2 \F_{\alpha \alpha} }\x^t \Chat^{-1} \Q^{\alpha} \Chat^{-1} \x,
\end{equation}
where
\begin{equation}
\hat{F}_{\alpha \alpha} \equiv \frac{1}{2} \textrm{tr} \left( \Chat^{-1} \Q^\alpha \Chat^{-1} \Q^\alpha \right),
\end{equation}
with the label ``loss" to foreshadow the fact that this will be an estimator with signal loss (i.e., a multiplicative bias of less than unity). We will now provide an explicit demonstration of this by modeling the estimated covariance as
\begin{equation}
\label{eq:ChatDef}
\Chat = (1-\eta) \C + \eta \x \x^t,
\end{equation}
where $\eta$ is parameter quantifying our success at estimating the true covariance matrix. If $\eta = 1$, our covariance estimate has perfectly modeled the true covariance and $\Chat = \C$. On the other hand, if $\eta =0$, then our covariance estimate is based purely on the one realization of the covariance that is our actual data, and we would expect a high level of overfitting and signal loss.

@@

This is shown in Figure \ref{fig:analytic_sig_loss} for $N=20$. As expected, the signal loss factor is unity for $\eta$ (perfect estimation of $\C$) and goes to zero as $\eta$ approaches $1$.

>>>>

To build our intuition into how a particular choice of weighting, namely inverse covariance weighting, can lead to signal loss, we use a toy model. We construct a simple dataset that contains 2-dimensional data (representing visibility data with $100$ time integrations and $20$ frequency channels). This model represents realistic dimensions of about an hour of PAPER data which might be used for a power spectrum analysis. We create a mock bright foreground signal, $\textbf{x}_{FG}$, as a complex sinusoid that varies in time and frequency, a simplistic but realistic representation of a single bright source. We also create a mock EoR signal, $\textbf{x}_{EoR}$, as a complex, Gaussian random signal. Our combined data vector is $\textbf{x} = \textbf{x}_{FG} + \textbf{x}_{EoR}$, to which we apply different weighting schemes. The three data vectors are shown in Figure \ref{fig:toy_sigloss1}.

@@

One choice for a weight matrix $\textbf{R}$ is an inverse covariance matrix. This type of weighting is attractive for power spectrum analyses because it is an aggressive way to down-weight foregrounds and yields the smallest possible error bars on a measurement (\citealt{tegmark_et_al1997a}; \citealt{bond_et_al1998}). The covariance matrix $\textbf{C}$, which in our case describes covariances between frequency channels, can be estimated in a variety of ways. Given perfect foreground, instrumental, and EoR models, we could form $\textbf{C}$ in a way that accurately describes our measured data. However, if our foreground model is flawed for example, our estimate of $\textbf{C}$ would not be successful at down-weighting them in the data.

@@

\noindent assuming $\langle\textbf{x}\rangle = 0$, where $\langle \rangle$ denotes a finite average over time $t$. The inverse covariance matrix is therefore $\textbf{R} = \hat{\textbf{C}}^{-1}$.

@@

\begin{figure*}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 1cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss3.png}
	\includegraphics[trim={1cm 0.2cm 0cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss4.png}
	\caption{Resulting power spectrum estimates for the toy model simulation described in Section \ref{sec:toymodel} --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We use inverse covariance weighting where $\textbf{C}$ is derived from the data (left), contrasted with projecting out the first eigenmode only (right). In the former case, signal loss arises from using information from all eigenmodes of $\hat{\textbf{C}}$. Because $\hat{\textbf{C}}$ is empirically estimated, its eigenvalues pose the risk of describing random fluctuations that happen to exist in the data realization but may not exist in the true covariance. Consequently, these modes are over-fitted and down-weighted, leading to signal loss. There is no signal loss when using only the zeroth eigenmode, since we are not using information from the weaker eigenmodes which $\hat{\textbf{C}}$ does not describe accurately.}
	\label{fig:toy_sigloss3}
\end{figure*}

As shown, our inverse covariance-weighted result successfully suppresses foregrounds. It is also evident that our result fails to recover the EoR signal --- it exhibits the correct shape, but the amplitude level is slightly low. This is evidence of signal loss. In order to understand the behavior of this result, we need to closely study our covariance matrix, $\hat{\textbf{C}}$, which is shown in Figure \ref{fig:toy_sigloss12}.

@@

If $\textbf{C}$ is computed from the data itself, it carries the risk of over-fitting information in the data and introducing a multiplicative bias to estimates of the signal. For a mathematical derivation of signal loss arising from a data-estimated covariance matrix, see Appendix \ref{sec:sigloss_appendix}. Here we will describe the origin of this signal loss intuitively.

It turns out that because we estimated $\textbf{C}$ from our data, its eigenspectrum offers insight into understanding how inverse covariance weighting affects our result. An eigenspectrum ranks the eigenvalues of a matrix from highest to lowest and can be thought of as a spectrum of weights that are given to each frequency mode in the data. In other words, the eigenvalues encode the strength of different shapes in the dataset. The eigenspectrum of the identity matrix $\textbf{I}$ is flat (all $1$'s) because it gives equal weighting to all modes. This is usually not the case for a covariance matrix, in which a sloped eigenspectrum means that modes are given different weights. When weighting data by an inverse covariance matrix, the modes with the highest eigenvalues are down-weighted the most.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss2.png}
	\caption{Eigenspectrum of $\hat{\textbf{C}}_{FG}$ (blue), $\hat{\textbf{C}}_{EoR}$ (red), and $\hat{\textbf{C}}_{FG+EoR}$ (green). The eigenspectrum of $\hat{\textbf{C}}_{FG}$ peaks at the zeroth eigenmode, due to the presence of only one sinusoid.}
	\label{fig:toy_sigloss2}
\end{figure}

If the true covariance matrix $\textbf{C}$ of our data was known, then every single eigenvalue of $\textbf{C}$ would be representative of real fluctuations in the data. However, when using an estimated $\hat{\textbf{C}}$ that is derived from one particular data realization, some of its eigenvalues may be describing random fluctuations that just happen to exist in the data realization. Said differently, shapes that may not exist (or have a weaker existence) in a true covariance may appear stronger in the estimated covariance. Hence, they will be down-weighted more than they should be.

In general, the strongest modes of $\hat{\textbf{C}}$ (highest eigenvalues) are more trustworthy than the weakest modes. Bright foregrounds usually dominate and would therefore exist in both $\textbf{C}$ and $\hat{\textbf{C}}$. This is demonstrated in the toy model by the successful suppression of the foreground mode, where our estimated covariance matrix identifies the sinusoid and assigns it the highest eigenvalue (the peak in Figure \ref{toy_sigloss2}.

The danger of an empirically estimated covariance matrix comes from not being able to describe weak eigenmodes accurately. The weak eigenmodes of $\hat{\textbf{C}}$ may characterize random noise fluctuations which will be down-weighted. This is what we call the `overfitting' of noise, which leads to signal loss.

Using what we've learned about the eigenspectrum, we can tweak it in a simple way to suppress foregrounds and yield zero signal loss. Namely, we can project out the zeroth eigenmode (i.e. zero out all eigenmodes except the first one), thereby down-weighting the foregrounds perfectly and nothing else. Hence, we are not using information from all the weaker eigenmodes which carry the risk of describing fluctuations that may not exist in the true covariance. Altering $\hat{\textbf{C}}$ in this way is one example of a regularization method, in which we are changing $\hat{\textbf{C}}$ in a way that flattens its eigenspectrum, making it more identical to that of $\textbf{I}$. The resulting power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. In this case we recover EoR, demonstrating that if we can model our foregrounds, we can down-weight them without signal loss. There are several other ways to regularize $\hat{\textbf{C}}$, and we will discuss some in Section \ref{sec:otherweight}.

@@

We have shown how signal loss can arise due to inaccurately characterizing weak eigenmodes with a data-estimated covariance. We will next show how this effect is exaggerated by reducing the total number of independent samples in a dataset.

@@

Because fringe-rate filtering is analogous to averaging in time, it comes at the cost of reducing the total number of independent samples in the data. To mimic this filter, we average every four time integrations of our toy model dataset together, yielding $25$ independent samples in time (Figure \ref{fig:toy_sigloss5}, left). We choose these numbers so that the total number of independent samples is similar to the number of frequency channels --- therefore, our matrices will be full rank.

The resulting eigenspectra (Figure \ref{fig:toy_sigloss5}, right), as compared to those in Figure \ref{fig:toy_sigloss2}, fall more steeply, especially for the last few eigenmodes. This is because we have fewer independent modes --- fewer shapes in the data --- so we do a worse job characterizing the weak eigenmodes. Therefore, there is a more dramatic difference in weighting between the modes and we end up down-weighting random noise-like fluctuations more severely than for the un-fringe-rate filtered case.

\begin{figure*}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss5.png}
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss6.png}
	\caption{Left: Our `fringe-rate filtered' (time-averaged) toy model dataset. We averaged every four samples together, yielding $25$ independent samples in time. Real parts are shown here. Right: Eigenspectrum of the covariance matrices of foregrounds only (blue), EoR only (red), and FG + EoR (green). The eigenspectrum of foregrounds only peaks at the zeroth eigenmode, a consequence of the presence of only one sinusoid. This spectrum is steeper than that in Figure \ref{fig:toy_sigloss2}, resulting in more dramatic difference in weighting between eigenmodes. Consequently, if all eigenmodes are used in weighting it is possible to down-weight random noisy modes more severely than for the un-fringe-rate filtered case.}
	\label{fig:toy_sigloss5}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss7.png}
	\caption{Resulting power spectrum estimate for the fringe-rate filtered toy model simulation --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We use inverse covariance weighting where $\textbf{C}$ is derived from the data. There is a larger amount of signal loss than for un-fringe-rate filtered data, a consequence of a steepened, weakly characterized eigenspectrum from using fewer independent modes in the data. This example shows that it is possible to down-weight modes more severely for the un-fringe-rate filtered since there is a more dramatic difference in weighting between modes. The power spectrum estimate is also noisier, since most of the information used in constructing the estimate came from only a few eigenmodes.}
	\label{fig:toy_sigloss7}
\end{figure}

The power spectrum results for inverse covariance weighted fringe-rate filtered data is shown in Figure \ref{fig:toy_sigloss7}. As expected, there is a much larger amount of signal loss for this time-averaged dataset. Additionally, because of the steepened eigenspectrum, most of our power spectrum information is coming from only the last few modes and as a result, it is a noisier estimate. This is evident by noticing that the green curve in Figure \ref{fig:toy_sigloss7} fails to trace the shape of the unweighted EoR power spectrum.

Using our toy model, we have seen that a sensitivity-driven analysis technique like fringe-rate filtering has trade-offs of signal loss and noisier estimates when using data-estimated covariance matrices. Longer integrations increase sensitivity but reduce the number of independent samples, resulting in weakly characterized, steep eigenspectra that can overfit noise greatly.

@@

In Section \ref{sec:toymodel} we showed one example of how altering $\hat{\textbf{C}}$ can make the difference between zero and some signal loss, if we can distinguish between real eigenmodes in a true covariance matrix from random fluctuation-induced eigenmodes in an estimated one. We will now use our toy model to describe several other ways to tailor $\hat{\textbf{C}}$ in order to minimize signal loss. We illustrate the resulting power spectra and eigenspectra for four different cases in Figures \ref{fig:toy_sigloss8} and \ref{fig:toy_sigloss14}.

@@

The second panel (top right) in Figure \ref{fig:toy_sigloss8} uses a regularization method of setting $\hat{\textbf{C}} = \hat{\textbf{C}} + \textbf{I}$. By adding the identity, we are weighting the diagonal elements of the matrix more heavily than those off-diagonal, thereby flattening out its eigenspectrum. If all modes are given similar weights, we avoid down-weighting artificial modes more than others.

The third panel (bottom left) in Figure \ref{fig:toy_sigloss8} flattens out the eigenspectrum of $\hat{\textbf{C}}$ a different way - by zeroing all but the first three eigenmodes. This means that the first three eigenmodes will be given weights, but the rest will be treated with similar weights. Again, flattening the eigenspectrum results in negligible signal loss. However, we do not perfectly recover the shape of EoR because we lost information when projecting out certain modes.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/data_errors.png}
	\caption{$2\sigma$ upper limit power spectrum estimates using PAPER-64 data and three different bootstrapping methods. The data is fringe-rate filtered and a weighting matrix of $\textbf{C}_{eff}^{-1}$ is used, where $\hat{\textbf{C}}_{eff} = \textbf{C}*\textbf{I}$, in order to minimize signal loss. Three different bootstrapping methods are shown: the black and red use methods where bootstrapping occurs on the baseline axis only (not the time axis), and they differ by how the baselines are randomly sampled. It is evident that sensitivity is maximized when baselines are sampled in a way that ensures mostly independent samples (only the last baseline spot is sampled randomly). The blue result bootstraps over both the baseline and time axes, and illustrates how errors can be under-estimated if sampling more values than independent ones (fringe-rate filtering reduces the number of independent samples).}
	\label{fig:data_errors}
\end{figure}

@@

\noindent Because of the nature of the lower triangular matrix, this window function has the property of preventing the leakage of foreground power from low $k$ to high $k$ modes. Specifically, we order the elements in $\textbf{F}$ in such a way so that power can leak from high $k$ modes to low $k$ modes, but not vice versa. Since most foreground power shows up at low $k$'s, this method ensures a window function that retains clean, noise-dominated measurements while minimizing the contamination of foreground bias.

@@

To mitigate crosstalk bias, which appears as a static bias in time, we apply a fringe-rate filter that suppresses fringe-rates of zero. Figure \ref{fig:frp} shows that the filter response is zero for such static signals. The effect of filtering out zero fringe-rates on power spectrum results is shown in \citet{ali_et_al2015}. Most notably, power spectrum detections exist at all $k$'s without crosstalk removal and these are detections that, depending on the power spectrum level, could be mistaken for EoR.

@@

\appendix
\section{A toy model for signal loss}
\label{sec:sigloss_appendix}

@@

The minimum-variance quadratic estimator $\hat{p}_\alpha$ for the $\alpha$th bandpower of the power spectrum is given by \acl{We need to decide whether the Q matrices have their Greek indices as upper or lower indices.} \acl{Also need to state in a footnote that for this section only we'll assume without loss of generality that the data are real}
\begin{equation}
\hat{p}_\alpha = \frac{1} {2 \F_{\alpha \alpha} }\x^t \C^{-1} \Q^{\alpha} \C^{-1} \x,
\end{equation}
where
\begin{equation}
F_{\alpha \alpha} \equiv \frac{1}{2} \textrm{tr} \left( \C^{-1} \Q^\alpha \C^{-1} \Q^\alpha \right)
\end{equation}
is the $\alpha$th diagonal element of the Fisher matrix. In our case, however, we do not have \emph{a priori} knowledge of the covariance matrix. Thus, we replace $\C$ with $\Chat$, its data-derived approximation. Our estimator then becomes
\begin{equation}
\label{eq:phatloss}
\hat{p}_\alpha^\textrm{loss} = \frac{1} {2 \F_{\alpha \alpha} }\x^t \Chat^{-1} \Q^{\alpha} \Chat^{-1} \x,
\end{equation}
where
\begin{equation}
\hat{F}_{\alpha \alpha} \equiv \frac{1}{2} \textrm{tr} \left( \Chat^{-1} \Q^\alpha \Chat^{-1} \Q^\alpha \right),
\end{equation}
with the label ``loss" to foreshadow the fact that this will be an estimator with signal loss (i.e., a multiplicative bias of less than unity). We will now provide an explicit demonstration of this by modeling the estimated covariance as
\begin{equation}
\label{eq:ChatDef}
\Chat = (1-\eta) \C + \eta \x \x^t,
\end{equation}
where $\eta$ is parameter quantifying our success at estimating the true covariance matrix. If $\eta = 0$, our covariance estimate has perfectly modeled the true covariance and $\Chat = \C$. On the other hand, if $\eta =1$, then our covariance estimate is based purely on the one realization of the covariance that is our actual data, and we would expect a high level of overfitting and signal loss.

@@

This is shown in Figure \ref{fig:analytic_sig_loss} for $N=20$. As expected, the signal loss factor is unity for $\eta=0$ (perfect estimation of $\C$) and goes to zero as $\eta$ approaches $1$.
